---
title: "HydroEnR-Welcome"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HydroEnR-Welcome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Welcome to the HydroEnR package
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Currently this is just a test page but more documentation will be added over time.

To load the package use;
```{r setup}
library(HydroEnR)
```

## QMED

Let's import some QMED data from NRFA, the site we'll use is 'St Johns Beck at Thirlmere Reservoir' - ID: 75001

For this we will require the rnrfa package.
```{r}
library(rnrfa)

data <- rnrfa::get_ts(id = 75001, type = 'amax-flow')
data
```

To find the QMED we will use the QMED function;
```{r}
qmed <- QMED(data)
qmed
```

Now we'll make a plot of the data
```{r}
plotQMED(data)
```

## L-moments

Within HydroEnR we can calculate the L-moments individually or as a grouped function.

To calculate the L1;
```{r}
L1(data)
```

You can also apply an urban adjustment. For this you can nest the function
```{r}
urbAdj(LCV(data), URBEXT2000 = 0.3)
```

Alternatively, you can use a piping operator from the magrittr library.
In this example we will set the DeUrb parameter to TRUE, normally set to FALSE.

```{r}
library(magrittr)
LCV(data) %>% 
  urbAdj(URBEXT2000 = 0.3, DeUrb = TRUE)
```

Using the Ls() function HydroEnR can bundle up all the L-moment calculations;
```{r}
Ls(data)
```

Urban adjustments can also be added with the URBEXT2000 and DeUrb parameters
```{r}
Ls(data, URBEXT2000 = 0.3, DeUrb = FALSE)
```

